---
title: "HW04 Part 1: Complete Chapter 3"
author: "William Zoughaib"
date: "r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options:
  chunk_output_type: inline 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```
### Chapter 3
### William Zoughaib

### Section 3.1
```{r}
library("tidyverse")
```

### Section 3.2 First steps
#running the gg plot and scatterplot data plot for mpg
```{r}
ggplot2::mpg
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))
```
#### Section 3.2 Questions
**1:** run ggplot(data=mpg):   
```{r}
ggplot(data=mpg)
```
#after running script, there is a blank graph with nothing on there.  I did not specify which x and y variables i needed
**2:** find dimension of mpg data
```{r}
dim(mpg)
```
# there are 234 rows of data and 11 columns.  So 234 x variables and 11 y variables
**3:** what does "drv" variable refer too?
```{r}
?mpg
```
#It refers to the type of drive train.  Meaning is it front wheel, rear wheel, or 4 wheel drive?

**4:** make a scatterplot of hwy v cyl
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=hwy, y=cyl))
```
**5:**make a scatterplot of class vs drv, why is it not useful:   
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=class, y=drv))
```
#after looking at the data, manufacturers build a variety of vehicles(classes) that will have different types of drivetrains.  An x variable, will be represented by possibly more than one y variable which takes away the x-y relationship


### Section 3.3 Aesthetic mappings
# using color = class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y= hwy, color =class))
```
#using size = class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy, size=class))
```
#using alpha=class & then shape=class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy,alpha=class))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,shape=class))
```
#use color=blue in aes() 
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy),color="blue")
```
#### Section 3.3 Questions
**1:** whats wrong with this code?  why is it red?
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy, color="blue"))
```
# i in fact made this mistake when i was reading the chapter itself.  To set an aesthetic, i need to type it outside of the aes() parantheses.  If not, ill just get the generic encoded color.
**2:** what variables are categorical, what variables are continuous: use ?mpg
```{r}
?mpg
```
# categorical:manufacturer, model, trans, drv, fl, class;    continuous:displ, year, cyl, cty, hwy
**3:** map a continuous variable( color for cyl, size=cyl, and shape=cyl)
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty, color=cyl))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty, size=cyl))
```
#attaching shape to a continuous variable results in an error.  ultimately, a continuous variable cannot be used with a shape.  Continuous variables can be used with color and size though.
**4:** map the continuous variable from above with the multiple aesthetics 
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty, color=cyl, size=cyl))
```
#i ran color=cyl and size=cyl and the above graph was shown.  Thus it appears that more than one aethetic for a continuous variable can be attributed to a scatterplot graph.
**5:** use stroke aesthetic and what shapes does it work with
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty,stroke=cyl))
```
#i do not like it as much as the size=cyl in the penultimate graph.  stroke=cyl works with circle shapes but it looks very blobby.  It shows a negative linear relationship but discerning a data pt is almost impossible.  
**6:** map an aesthetic to something other than a variable name such as color=displ<5
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty,color=displ<5))
```
#this is pretty cool as adding it to the code chunk then provides boolean values of TRUE, FALSE.  If i chose to do this, it would help decrease the required data range


###Section 3.5: Facets
#use facet wrap on mpg data
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_wrap(~class, nrow=2)
```
#run facet grid on mpg data
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_grid(drv~cyl)
```
#run a period instead of a variable name
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_grid(.~cyl)
```
####3.5.1 Exercises
**1:** What happens if a you facet on a continuous variable
#it will run and only show the continuous variable above displ versus the hwy mpg
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_wrap(~year,nrow=2)
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_wrap(~displ,nrow=2)
```
#i have run two different code chunks with continuous variables and it still works
**2:** what do the empty cells in code chunk w/ facet_grid(drv~cyl) mean; how does it relate to the following code chunk
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_grid(drv~cyl)
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=drv, y=cyl))
```
#it is telling us that there are 0 cars with 4 cylinders and rear wheel drive; it is also telling us there are 0 cars with either 5 cylinder & 4 wheel drive, 5 cylinder & rear wheel drive
**3:**what plots do the following code chunks make.  what does the . do
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))+
  facet_grid(drv~.)
```
#the . eliminates column data and only shows the extra row data represented by drive train along with the hwy and displ variables
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))
facet_grid(. ~ cyl)
```
#this plot looks very similar to ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy)) code chunk if it isnt the same.  Apparently my continuous variable cyl is not shown.  
**4:** take the first faceted plot we did above and determine the advantages and disadvantages of it.  how might the balance change
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))+
  facet_wrap(~class,nrow=2)
```
#Instead of trying to determine the different colors using color aesthetics, facet_wrap simplifies the different classes into their own scatterplot.  It is easier to read when trying to highlight a specific portion of the mpg dataset.  But the color aesthetic has it uses when looking at something on a grand scale where we are just trying to determine the overall linear relationship.  You cannot determine the overall linear relationship using facet_wrap ie 2seater and minivan classes.  Larger dataset would be more useful with the color aesthetic as we can look at the overall picture better than with the facet_wrap.  If we wanted to of course obtain a specific part of the larger data set, then we can use facet_wrap
**5:** read ?facet_wrap.  what does nrow and ncol do.  what other options are there to control the layout.  Why doesnt facet_grid have nrow and ncol arguments
```{r}
?facet_wrap
```
#it offers a better screen space than facet_grid.  we can compare each categorical variable to each other much better with facet_wrap. nrow= number of desired rows, ncol = number of desired columns.  These include as.table, switch, dir, shrink, scales, drop, strip.position)
**6:** why should you put the variable with more unique levels in the columns with facet_grid?
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))+
  facet_grid(drv~cyl)
```
# i do not know why?  is it because people like reading horizontally?  More columns helps the eyes when reading from left to right...also do not have to scroll as much.  So the column variable should have more unique variables than the row variable.  


###3.6 Geomteric Objects
#use geom_smooth to plot mpg data
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy))
```
#add linetype to drv with above code chunk
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,linetype=drv))
```
#add group=drv to original geom_smooth plot
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,group=drv))
```
#repeat above code chunk but adding color=drv and show.legend=FALSE
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,color=drv),show.legend = FALSE)
```
#add geom_point and geom_smooth to mpg scatter plot of displ v hwy
```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  geom_smooth(mapping = aes(x=displ,y=hwy))
```
#to save typing and have two geom plots, add mapping =aes(x=displ,y=hwy) to ggplot(data=mpg,) code chunk
```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy))+
  geom_point()+
  geom_smooth()
```
#apply color aesthetic to geom_point only while leaving geom_smooth normal
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth()
```
#repeat above code chunk but adding in geom_smooth by selecting subcompact cars for class and using mpg to filter
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)
```
####3.6.1 Exercises
**1:** determine geom_ for line chart, boxplot,histogram,area chart
# respectively the answers are geom_line, geom_boxplot, geom_histogram, geom_area
**2:**
#It will show points on the graph but in the geom_smooth, the se will override the global data and just show curved lines
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy,color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
```
**3:** what does show.legend=FALSE do? what happens if you remove it.  copy original data chunk that had show.legend but remove it
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,color=drv))
```
#show.legend=FALSE removes the legend key to the left of the smooth data plot.  By removingshow.legend=false, you restore the original argument that show.legend=TRUE thus the legend will appear.  It was used to show the complete functionality of the geom_smooth function.  When adding the color aesthetic to a variable, it is useful to know which color matches what attribute of that specific variable thus a legend is usually required
**4:** what does se do in geom_smooth
#it overwrites the ggplot data overall data and only selects the geom_smooth data we wish to see.  
**5:**will these two graphs look different? 
#No, they will be the same. when plugging in the mappings to ggplot(data=mpg,mapping....) it overwrites what mapping of the data we need in the geom_smooth and geom_point chunks.  by plugging in mapping into ggplot (we have to use the same x,y variables), we avoid needless extra typing
**6:** 
#1:upper left
```{r}
ggplot(data=mpg, mapping = aes(x=displ,y=hwy))+
  geom_point()+
  geom_smooth(se=FALSE)
```
#2:upper right
```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,group=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
```
#3:middle left
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy,color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
```
#4:middle right
```{r}
ggplot(data=mpg, mapping = aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=drv))+
  geom_smooth(se=FALSE)
```
#5:lower left
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=drv))+
  geom_smooth(mapping=aes(linetype=drv),se=FALSE)
```
#6:lower right
```{r}
?geom_point
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,fill=drv),shape=21,stroke=2,size=3,color="white")
```


###Section 3.8 Position Adjustments
#run geom_bar and provide color aesthetic and use fill=cut
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,color=cut))
```
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=cut))
```
#run fill=clarity with above code chunk by replacing fill=cut
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut,fill=clarity))
```
#run alpha=1/5 with position="identity" in geom_bar chunk by using ggplot(data=diamonds,mapping=aes(x=cut,fill=clarity))
```{r}
ggplot(data=diamonds,mapping=aes(x=cut,fill=clarity))+
  geom_bar(alpha=1/5,position="identity")
```
#repeat above step but replace alpha=1/5 with fill=NA and color=clarity
```{r}
ggplot(data=diamonds,mapping=aes(x=cut,color=clarity))+
  geom_bar(fill=NA,position="identity")
```
#run position="fill" to place overlapping objects right next to each other
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="fill")
```
#run position="dodge" by replacing position="fill" as in the above code chunk which runs overlapping objects right next to each other
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="dodge")
```
#in original mpg plot of displ v hwy mpg data, only 109 d.p. were shown.  Use position="jitter" to show all data points
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy),position="jitter")
```
####3.8.1 Exercises
**1:** run the plot and what is the problem with it?  TO fix it, run geom_jitter
```{r}
ggplot(data=mpg,mapping=aes(x=cty,y=hwy))+
  geom_point()
```
```{r}
ggplot(data=mpg)+
  geom_jitter(mapping=aes(x=cty,y=hwy))
```
#It is missing alot of the data if we kept the original plot.  Using geom_jitter mapping, we can see all the data points for the 234 cars
**2:** what parameters to geom_jitter control jittering amount
```{r}
?geom_jitter
```
#it overwrites the tidy plot data if we just did geom_point(mapping=aes(x=cty,y=hwy)) and shows all points.  width, height, position are the parameters
**3:**Compare and contrast geom_jitter with geom_count
```{r}
ggplot(data=mpg)+
  geom_jitter(mapping=aes(x=cty,y=hwy))
```
```{r}
ggplot(data=mpg)+
  geom_count(mapping=aes(x=cty,y=hwy))
```
```{r}
?geom_count
```
# geom_count has a legend attached.  It also has dots that vary in size according the data point that falls there (how many objects fit the x,y data point, how many observations are there at each location).  Geom_count and geom_jitter are similar in that they have a positive linear relationship. They can both be used with overplotting and they both can use aesthetics such as color, fill, shape, etc. 
**4:**what is the default position adjustment for geom_boxplot.  show the mpg dataset that portrays it by running position=: "dodge" versus "dodge2"
```{r}
?geom_boxplot
```
# the default position is "dodge2"
```{r}
ggplot(data=mpg)+
  geom_boxplot(mapping=aes(x=drv,y=hwy),position="dodge")
```
```{r}
ggplot(data=mpg)+
  geom_boxplot(mapping=aes(x=drv,y=hwy),position="dodge2")
```
#here is the boxplot for it


###Section 3.9: Coordinate Systems
#run geom_boxplot for class v hwy mpg data; then run with coord_flip()
```{r}
ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()
```
```{r}
ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()+
  coord_flip()
```
#cont. reading


###3.10:The layered grammar of graphics
#cont. reading

