---
title: "HW04 Part 1: Complete Chapter 3"
author: "William Zoughaib"
date: "r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options:
  chunk_output_type: inline 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```
### Chapter 3
### William Zoughaib

### Section 3.1
```{r}
library("tidyverse")
```

### Section 3.2 First steps
#running the gg plot and scatterplot data plot for mpg
```{r}
ggplot2::mpg
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))
```
#### Section 3.2 Questions
**1:** run ggplot(data=mpg):   
```{r}
ggplot(data=mpg)
```
#after running script, there is a blank graph with nothing on there.  I did not specify which x and y variables i needed
**2:** find dimension of mpg data
```{r}
dim(mpg)
```
# there are 234 rows of data and 11 columns.  So 234 x variables and 11 y variables
**3:** what does "drv" variable refer too?
```{r}
?mpg
```
#It refers to the type of drive train.  Meaning is it front wheel, rear wheel, or 4 wheel drive?

**4:** make a scatterplot of hwy v cyl
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=hwy, y=cyl))
```
**5:**make a scatterplot of class vs drv, why is it not useful:   
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=class, y=drv))
```
#after looking at the data, manufacturers build a variety of vehicles(classes) that will have different types of drivetrains.  An x variable, will be represented by possibly more than one y variable which takes away the x-y relationship


### Section 3.3 Aesthetic mappings
# using color = class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y= hwy, color =class))
```
#using size = class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy, size=class))
```
#using alpha=class & then shape=class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy,alpha=class))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,shape=class))
```

#### Section 3.4 Questions









Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
