---
title: "HW04 Part 1: Complete Chapter 3"
author: "William Zoughaib"
date: "r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options:
  chunk_output_type: inline 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```
### Chapter 3
### William Zoughaib

### Section 3.1
```{r}
library("tidyverse")
```

### Section 3.2 First steps
#running the gg plot and scatterplot data plot for mpg
```{r}
ggplot2::mpg
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))
```
#### Section 3.2 Questions
**1:** run ggplot(data=mpg):   
```{r}
ggplot(data=mpg)
```
#after running script, there is a blank graph with nothing on there.  I did not specify which x and y variables i needed
**2:** find dimension of mpg data
```{r}
dim(mpg)
```
# there are 234 rows of data and 11 columns.  So 234 x variables and 11 y variables
**3:** what does "drv" variable refer too?
```{r}
?mpg
```
#It refers to the type of drive train.  Meaning is it front wheel, rear wheel, or 4 wheel drive?

**4:** make a scatterplot of hwy v cyl
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=hwy, y=cyl))
```
**5:**make a scatterplot of class vs drv, why is it not useful:   
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=class, y=drv))
```
#after looking at the data, manufacturers build a variety of vehicles(classes) that will have different types of drivetrains.  An x variable, will be represented by possibly more than one y variable which takes away the x-y relationship


### Section 3.3 Aesthetic mappings
# using color = class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y= hwy, color =class))
```
#using size = class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy, size=class))
```
#using alpha=class & then shape=class
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy,alpha=class))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,shape=class))
```
#use color=blue in aes() 
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy),color="blue")
```
#### Section 3.3 Questions
**1:** whats wrong with this code?  why is it red?
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy, color="blue"))
```
# i in fact made this mistake when i was reading the chapter itself.  To set an aesthetic, i need to type it outside of the aes() parantheses.  If not, ill just get the generic encoded color.
**2:** what variables are categorical, what variables are continuous: use ?mpg
```{r}
?mpg
```
# categorical:manufacturer, model, trans, drv, fl, class;    continuous:displ, year, cyl, cty, hwy
**3:** map a continuous variable( color for cyl, size=cyl, and shape=cyl)
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty, color=cyl))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty, size=cyl))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty,shape=cyl))
```
#attaching shape to a continuous variable results in an error.  ultimately, a continuous variable cannot be used with a shape.  Continuous variables can be used with color and size though.
**4:** map the continuous variable from above with the multiple aesthetics 
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty, color=cyl, size=cyl))
```
#i ran color=cyl and size=cyl and the above graph was shown.  Thus it appears that more than one aethetic for a continuous variable can be attributed to a scatterplot graph.
**5:** use stroke aesthetic and what shapes does it work with
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty,stroke=cyl))
```
#i do not like it as much as the size=cyl in the penultimate graph.  stroke=cyl works with circle shapes but it looks very blobby.  It shows a negative linear relationship but discerning a data pt is almost impossible.  
**6:** map an aesthetic to something other than a variable name such as color=displ<5
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty,color=displ<5))
```
#this is pretty cool as adding it to the code chunk then provides boolean values of TRUE, FALSE.  If i chose to do this, it would help decrease the required data range


###Section 3.5: Facets
#use facet wrap on mpg data
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=cty))+
  facet_wrap(~class, nrow=2)
```








