---
title: "HW 12: Data Visualization 1 cont"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
library(tidyverse)
library(here)
library(lubridate)
```
***
#### Part 1: Play Date (copied and pasted)
```{r}
birth_date <- "05/14/1998"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Let's also remember a dream
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 28"
```
#### Convert each date to POSIXct format, using lubridate functions, saving each to a unique variable.  (I guess I am adding a time of day?)
```{r}
birth_time <- mdy(birth_date)
birth_time
# I really dont know when I was born.  
anning_time <- ymd(anning)
anning_time
# I would hunt for fossils in the afternoon
ball_time <- mdy(ball)
ball_time
# I am most preductive with science in the morning
earle_time <- mdy(earle)
earle_time
# Lunch time
franklin_time <- dmy(franklin)
franklin_time
# nasty little hobbitses
lederberg_time <- mdy(lederberg)
lederberg_time
# oof, i ran out
mcclintock_time <- dmy(mcclintock)
mcclintock_time
# This time is confusing
mlk_birth_time <- mdy(mlk_birth)
mlk_birth_time
mlk_dream_time <- ymd(mlk_dream)
mlk_dream_time
# Pertinent
```
#### Check yours dates to be sure they are correct. If you spot any errors, correct them now using the proper lubridate function(s).
```{r}
mlk_birth.1 <- "01/15/1929"
mlk_birth_time.1 <- mdy(mlk_birth.1)
mlk_birth_time.1
# edited birth year from 2029 to 1929.  
```
#### What day of the week were you born on? Show the full day name, not the abbreviated name or the day number.
```{r}
wday(birth_time,label=TRUE,abbr=FALSE)
# Thursday was the day I was born on.  
```
####What day of the week will your birthday be on this year? Use lubridate’s update function. You must figure out how to do it entirely by code for full points (no hard-coding). In other words, you code should work just as well next year or five years in the future, without editing.`
```{r}
find_my_day_birth <- function(year,month,day){
  update(birth_time,year,month,day)}
  wday(find_my_day_birth(year=2020,month=5,day=14),label=TRUE,abbr=FALSE)
# I am really confused by this.  This is the only way I could get it to work without errors. My birthday fell on a Thursday this year.   
```
```{r}
wday(find_my_day_birth(year=2024,month=5,day=14),label=TRUE,abbr=FALSE)
  # THis still works
```
```{r}
wday(find_my_day_birth(year=2020,month=10,day=5),label=TRUE,abbr=FALSE)
  # using Mom's birthdate, it still works.  
```
```{r}
wday(find_my_day_birth(year=2020,month=4,day=26),label=TRUE,abbr=FALSE)
# My sisters still works.  Sorry for all these extra code chunks.  
```
#### What day of the week would your birthday be on in the Year 2525, if Man is still alive? How old would you be? Here you may hard-code 2525. Show the abbreviated day of the week.
```{r}
wday(update(birth_time,year=2525),label=TRUE)
# Hard code
```
```{r}
wday(find_my_day_birth(year=2525,month=5,day=14),label=TRUE)
# Using my find_my_day_birth function
```
#### How many days was it between the birth of Dr. Elizabeth Garrett Anderson and Alice Augusta Ball?
