---
title: "HW 08: More Data Wrangling"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.



## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

**You do not answer any of the questions this week.** Just read and run all of the examples as noted below. Pay attention to the code you enter for the examples. You will be applying the information in future assignments.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 
#### Running mutate() 
```{r}
flights_sml <- select(flights,year:day,ends_with("delay"),distance,air_time)
mutate(flights_sml,gain=dep_delay-arr_delay,speed=distance/air_time*60)
# by running this code, i added year:day columns, dep_delay,arr_delay, distance & air_time.  By using mutate function, i then created columns gain and speed.  
```
```{r}
mutate(flights_sml,gain=dep_delay-arr_delay,hours=air_time/60,gain_per_hour=gain/hours)
# By using the flights_sml vector, i then confirmed gain, determined hours of flight.  I could now use these columns to determine gain per hour.  
```
#### Running transmute ()
```{r}
transmute(flights,gain=dep_delay-arr_delay,hours=air_time/60,gain_per_hour=gain/hours)
# transmute function only keeps the columns I want created.
```
```{r}
transmute(flights,dep_time,hour=dep_time%/%100,minute=dep_time%%100)
# By using %/% and %% I was able to break down the dep_time into the specific number of hours and the specific number of minutes
```




### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.
#### Using summarise()
```{r}
summarise(flights,delay=mean(dep_delay,na.rm=TRUE))
# obtaining mean of dep_delay
```
```{r}
by_day <- group_by(flights,year,month,day)
summarise(by_day,delay=mean(dep_delay,na.rm=TRUE),.groups = 'drop')
# obtaining the mean of dep_delay on each day for 365 days
```
#### Using group_by and summarise together to make better data interpretation
```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
   count = n(),
   dist = mean(distance, na.rm = TRUE),
   delay = mean(arr_delay, na.rm = TRUE),.groups='drop')
delay <- filter(delay,count>20,dest!="HNL")
ggplot(data=delay,mapping=aes(x=dist,y=delay))+
  geom_point(aes(size=count),alpha=1/3)+
  geom_smooth(se=FALSE)
# Apparently there is an increase in delay as flights get longer around up to 750 miles.  There is a decrease in delay though past the 750 mile point.  
```
```{r}
delay
```

#### Using pipes to do what I did above
```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count=n(),
    dist=mean(distance,na.rm=TRUE),
    delay=mean(arr_delay,na.rm=TRUE),.groups='drop') %>%
filter(count>20,dest!="HNL")
```
```{r}
delays
# Same thing as previous code chunk but word delays instead of delay.  %>% is so much easier.  
```
#### Removing na.rm argument
```{r}
flights %>% 
  group_by(year,month,day) %>% 
  summarise(mean=mean(dep_delay))
# mean column is missing all the critical data!!!
```
```{r}
flights %>% 
  group_by(year,month,day) %>% 
  summarise(mean=mean(dep_delay,na.rm=TRUE))
# much nicer now when using na.rm argument
```
####Removing missing values which equal cancelled flights
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay),!is.na(arr_delay))
not_cancelled %>% 
  group_by(year,month,day) %>% 
  summarise(mean=mean(dep_delay),.groups='drop')
# By removing cancelled flights, mean decreased.  Using this code chunk, it helps remove outlier data
```
#### Using counts and creating delayss.  Use geom_freqpoly
```{r}
delayss <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(delay=mean(arr_delay),.groups='drop')
ggplot(data=delayss,mapping=aes(x=delay))+
  geom_freqpoly(binwidth=10)
# Interestingly this is a new geom_function and it shows that there are delays up to 330 minutes
```
#### use scatterplot now
```{r}
delayss <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay=mean(arr_delay,na.rm=TRUE),
    n=n(),.groups='drop')
ggplot(data=delayss,mapping=aes(x=n,y=delay))+
  geom_point(alpha=1/10)
# Greatest variation is where the flights are fewer in number
```
#### Filtering above plot
```{r}
delayss %>% 
  filter(n>25) %>% 
  ggplot(mapping=aes(x=n,y=delay))+
  geom_point(alpha=1/10)
# removing flights under 25
```
#### **I prefer Cycling**

### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

