---
title: "HW07 Data Importing and Wrangling"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("here")
```
```{r}
library(ggplot2)
```

```{r}
library(tidyverse)
library(here)
```
```{r}
library(ggplot2)
```


###Assignment 1: **Limpets**
```{r}
limpets_data <- read_csv(here("data","limpets.csv"),col_types="ncc")
```
```{r}
limpets_data
```
####Are the data tidy?
```{r}
as_tibble(limpets_data)
# tidying the data
# data is the same as previous code chunk so data is tidy
```
####Set col_types argument to numeric,character,and character
```{r}
# col_types have already been set to numeric, character,character
```


####Make sites column an ordered factor
```{r}
?factor
```
```{r}
limpets_data$Site <- factor(limpets_data$Site,levels=c("PBL","KNRM","VBG","WP","PF","DP","SIO","CTZ1","CTZ2","CNM"),ordered=TRUE)
```
```{r}
limpets_data
```



####Create a boxplot for limpets data
```{r}
?labs
```
```{r}
ggplot(data=limpets_data)+
  geom_boxplot(mapping=aes(x=Length,y=Site,color=Protected))+labs(x="Length (mm)")+
  labs()
```
#### DP and WP are the only two boxplots with outliers.  It can be observed that protected sites have larger limpets that unprotected sites.  


###Assignment 2: **Roseate Terns**
```{r}
library("tidyverse")
```
```{r}
?read_tsv
```
```{r}
terns_data <- read_tsv(here("data","roseate_terns.txt"))
# using read_tsv instead of read_csv
```
```{r}
terns_data
```


####Tidy the data
```{r}
tidy_terns <- terns_data %>% pivot_longer(c(AR:JC),names_to="Site",values_to="Population")
# remember to use the %>% function as it is very very useful.  Also, syntactic words do not get ` `.  
```
```{r}
tidy_terns
```


####Filter out missing Counts
```{r}
?filter
```
```{r}
terns_filtered <- filter(drop_na(tidy_terns))
```
```{r}
terns_filtered
```



####Create a Line Plot of population size over time
```{r}
ggplot(data=terns_filtered)+
  geom_line(mapping=aes(x=Population,y=year,color=Site))+
  labs(y="Year")
```
#### Sites AR,BC,JC obviously increased over time.  Site JR and BR had obvious sharp decreases in population but it appears that there is a population resurge.  Some of the others are really hard to read.  But in one example, (AR), it appears that it was introduced into that site around 1995 whie otheres were introduced earlier.  Some species may have been wiped out as well.  This would contribute to breaks.  



###Assignment 3: Blacklip Abalone
#### Chunk 1: import abolone data, remove first column, and make a boxplot of height differences
```{r}
abalone_data <- read_csv(here("data","abalone.csv"))
select(abalone_data,-(X1))
ggplot(data=abalone_data)+
  geom_boxplot(mapping=aes(x=Height,y=Type))

# using read_csv, then selecting against X1 column, then creating a boxplot by using Height v Type (originally used height v type)
```
#### Chunk 2: create a scatterplot to see if height correlates with rings
```{r}
ggplot(data=abalone_data)+
  geom_point(mapping=aes(x=Rings,y=Height))
# using geom_point to graph RIngs v height
```
#### There is a general increase in height as there are more rings.  But the two outliers are not explained by it as the rings are <10.  


#### Chunk 3: Filter out the outliers.  Remove the two large indivuals.  Change Type to an ordered factor.  Immatures are first.  Redo the scatterplot
```{r}
abalone_clean <- filter(abalone_data,Height<=0.4)
factor(abalone_clean$Type,levels=c('I','F','M'))
ggplot(data=abalone_clean)+
  geom_point(mapping=aes(x=Height,y=Rings,color=Type))
```
#### Immature ones are found in the lower right area thus they are smaller in height and have fewer rings.  Makes perfect sense.  Males and females appear to be equally dispersed for the most part.  There are a couple outliers where a few females and males are smaller than immature ones.  The same goes for immature ones where they are found among large male and females.  


#### Chunk 4: Select the individuals with zero height and assign them an NA value.  Redo scatterplot
```{r}
abalone_clean[abalone_clean$Height == 0, 'Height'] 
ggplot(data=abalone_clean)+
  geom_point(mapping = aes(x=Height,y=Rings,color=Type))
# I think this is the correct way
```
```{r}
is.na <- filter(abalone_data,Height==0)
ggplot(data=abalone_clean)+
  geom_point(mapping=aes(x=Height,y=Rings,color=Type))
# this is the same scatterplot as above except that I filtered it out.  
```
  

#### Chunk 5: Create two different scatterplots of continuous variables.  Use color and/or shape.
```{r}
abalone_clean
```
```{r}
ggplot(data=abalone_clean)+
  geom_point(mapping=aes(x=LongestShell,y=ShellWeight,color=Type,shape=Type))
```
#### In here we have a curve in the data.  Immatures will have the lower ShellWeight v. LongestShell versus their Female and Male counterparts.  
```{r}
ggplot(data=abalone_clean)+
  geom_point(mapping=aes(x=Rings,y=Diameter,color=Type,shape=Type))
```
####Interestingly, Diameter is greatest for male and female at around 10-12 rings.  Then there is a slight decrease in diameter as rings are added.  Again immatures will have a smaller ratio of Rings to Diameter than the adult counterparts.  


###Assignment 4: Darters
```{r}
?read_fwf
```
```{r}
library(tidyverse)
```
####Load darters data.  Assign Column names. Make Riffle an unordered factor while Maj_Type is an ordered factor
```{r}
darters_data <- read_fwf(here("data", "darter.txt"), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2)), comment = '#')
darters_named <- darters_data %>% rename(ID=X1,Species=X2,Riffle=X3,Age=X4,Sex=X5,Length=X6,Depth=X7,Velocity=X8,Temperature=X9,Maj_Sub=X10,Min_Sub=X11,Maj_Type=X12,Min_Type=X13)
darters_named$Riffle <- factor(darters_named$Riffle,levels=c('1','2'))
darters_named$Maj_Type <- factor(darters_named$Maj_Type,levels=c('s','fg','sg','lg','c'),ordered=TRUE)
```
```{r}
darters_data
# without any of the above corrections.
```
```{r}
darters_named
# checking to see that column names were changed. Ordered and factor were added to Maj_Type and Riffle respectively
```






#### Using pipe, filter out "zonale" & "tetrazonum".  Remove mintype and minsub columns.  Rename Maj_Sub and Maj_Type to Major_substrate and Major_type respectively.  Arrange data by id.  
```{r}
darters_tidy <- darters_named %>% 
  filter(Species=="blennioides"|Species=="caeruleum"|Species=="spectabile") %>% 
  rename(
    Major_Substrate = Maj_Sub,
    Major_Type = Maj_Type) 
select(darters_tidy,-(c(Min_Type,Min_Sub))) %>% 
  arrange(darters_tidy,ID) %>% 
  drop_na()
# I think i did it?  I now have 119 rows of data only corresponding to the IDs.  Then I did drop_na and I removed one extra row.   
```


#### Graph 3 plots using Riffles.  Plot 1: depth v length (Species= color & shape).  Plot 2: Boxplot of species v length.  Which riffle shows the greatest number of outliers.  Plot 3Boxplot of length for major substrate types for each species for each riffle.  Six plots in one!
```{r}
ggplot(darters_tidy)+
  geom_point(mapping=aes(x=Depth,y=Length,color=Species,shape=Species))+
  facet_wrap(~Riffle,ncol=2)
####There are fewer blennioides in the first riffle compared to the second.  The first riffle has greater depth overall while the second riffle has long length.  
```


```{r}
ggplot(data=darters_tidy)+
  geom_boxplot(mapping=aes(x=Species,y=Length))+
  facet_wrap(~Riffle,ncol=2)
# riffle 2 has more outliers than riffle 1
```
```{r}
ggplot(data=darters_tidy)+
  geom_boxplot(mapping=aes(x=Major_Substrate,y=Length))+
  facet_grid(Species~Riffle)
# this looks right but I do not understand the triangle warning sign?
```
```{r}
ggplot(data=darters_tidy)+
  geom_boxplot(mapping=aes(x=Major_Substrate,y=Length))+
  facet_grid(Riffle~Species)
# This looks nicer as I can compare each species to what riffle it was in.  Riffle 2 again shows more outliers.  
```



###THE END




