---
title: "HW 09: Wrangling 2"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(here)
library(knitr)
library(smatr)
```
```{r}
source(here("scripts","my_functions.R"))
```


### **Darters in Riffles**
***
#### Import data, filter out tetrazonum & zonale
```{r}
darters_data <- read_csv(here("data","darters.csv"))
darters_filtered <- darters_data %>% 
  filter(species=="blennioides"|species=="caeruleum"|species=="spectabile")
darters_filtered
```
***
#### Wrangle data.  Use Mutate to change Riffle values, centimeter values, create total_substrate column, change f & m to female and male
```{r}
mut_darters <- mutate(darters_filtered,total_substrate=majsub+minsub) %>% 
  mutate(riffle = ifelse(riffle == "1", 
                             "Riffle 1", 
                             "Riffle 2")) %>% 
  mutate(length=length*10) %>% 
  mutate(sex=ifelse(sex=="m","Male","Female"))
# all values under total_substrate = 1
```
***
#### Summarize Data: use sum_darters
```{r}
sum_darters <- group_by(mut_darters,species,riffle)   
darters_sum <- summarise(sum_darters,
  mean_length=mean(length,na.rm=TRUE), 
  mean_depth=mean(depth,na.rm=TRUE),
  mean_velocity=mean(velocity,na.rm=TRUE),.groups='drop')
darters_sum
# narrowed it down to the five columns i wanted which are species, riffle, mean_length,mean_depth.  This was tough.
```
#### Run kable on newly created darters_sum
```{r}
kable(darters_sum)
# This looks cool
```
***
#### Graph Data
####Plot 1
***
##### Run stat_summary
```{r}
ggplot(data=sum_darters) + 
  stat_summary(
     mapping = aes(x=species,y=length),
     fun.ymin = min,
     fun.ymax = max,
     fun.y = median,
     na.rm=TRUE)
```
***
##### RUn stat_summary and geom_point (use darters_sum)
```{r}
ggplot(data=darters_sum,mapping=aes(x=species,y=mean_length))+
  geom_point(shape=15,color="red")+
  stat_summary(mapping = aes(x=species,y=length),
     fun.ymin = min,
     fun.ymax = max,
     fun.y = median,
     data=sum_darters,
     na.rm = TRUE)
# I had to use ?stat_summary to find out how to overlay mean _length over length.  I think this is how you do it?  You have to determine what the data is somehow.  I also added color to the square shape as it will be easier to notice
```
***
##### Rearrange placements of stat summary and geom_point to confirm plot
```{r}
ggplot(data=sum_darters) + 
  stat_summary(
     mapping = aes(x=species,y=length),
     fun.ymin = min,
     fun.ymax = max,
     fun.y = median,
     na.rm=TRUE)+
  geom_point(mapping=aes(x=species,y=mean_length),data=darters_sum,shape=15,color="red")
# Swapping the sum_darters with darters_sum and changing plot functions still results in the same thing.
```
***
##### Run coord_flip() on the stat_summary with geom_point overlay.  Run facet_wrap to seperate the two riffles and add labs() to include measurement
```{r}
ggplot(data=sum_darters) + 
  stat_summary(
     mapping = aes(x=species,y=length),
     fun.ymin = min,
     fun.ymax = max,
     fun.y = median,
     na.rm=TRUE)+
  geom_point(mapping=aes(x=species,y=mean_length),data=darters_sum,shape=15,color="red")+
  facet_wrap(~riffle,ncol=2)+
  labs(x=NULL)+
  labs(y="length(mm)")+
  coord_flip()
# I can observe the two riffles now with labels changed.  There is greater variance in Riffle 2 than in Riffle 1
```
***
#### Plot 2.  Change color,add stat_summary, another stat_summary, facet,label,coord_flip()
```{r}
ggplot(data=sum_darters) + 
  stat_summary(
     mapping = aes(x=species,y=length),
     fun.ymin = min,
     fun.ymax = max,
     geom="point",
     size=2.2,
     fun.y = median,
     na.rm=TRUE
     )+
  stat_summary(
    mapping=aes(x=species,y=length),
    fun.ymin = min,
    fun.ymax = max,
    size=3,
    color="maroon",
    fun.y=mean,
    na.rm=TRUE
  )+
  geom_point(mapping=aes(x=species,y=mean_length),data=darters_sum,color="grey66")+
  facet_wrap(~riffle,ncol=2)+
  labs(x=NULL,y="length(mm)")+
  coord_flip()
# I think I followed all the steps.  "Grey66" is in the geom_point, first stat_summary has median,"point", & 2.2.  Second stat_summary has mean,3, and maroon.  It is just not what I expected.  
```







### Stress and corticosterones in birds
***
#### Import Data
```{r}
feathers_data <- read_csv(here("data","feathers.csv"),locale = locale(decimal_mark = ",")) %>% 
  rename(capture_date="Capture Date",day=Date,site=Site,urban_score="Urba Score",cortF="CortF ng/mg",cortB="CortF ng/mm", sex=Sex,smi=SMI,tarsus=Tarsus,mass=Mass,basal_cort=Bas_Cort,stress_cort=Stress_Cort)
feathers_data
# imported data, then changed decimal marks to "." instead of ",".  I then changed column names.  I needed to remember to do "around spaced words" to change column name.  Saved data as feathers_data.  
```

***
#### Calculate _b_ slope of the regression. Comments provide directions
```{r}
mean_tarsus <- mean(tarsus,na.rm = TRUE)
mean_tarsus
# find mean tarsus length and store in mean_tarsus.
```
```{r}
mut_feathers <- mutate(feathers_data,log_mass=log(mass),log_tarsus=log(tarsus))
mut_feathers
# Mutate to log transform mass and log transform tarsus.  Use log_mass and log_tarsus.  
```
```{r}
major_axis <- sma(log_mass~log_tarsus,data=mut_feathers)
summary(major_axis)
# create major_axis using sma(log_mass~log_tarsus,data=mut_darters)...**DO NOT USE -** or you will have issues!!!!!!  Then apply summary function
```
```{r}
coef(major_axis)
major_axis_coef <- coef(major_axis)
# simply showing the estimate values for elevation and slope
b <- major_axis_coef['slope']
b
# storing b value for slope estimate
```
#### Now that we have b, use SMI function
```{r}
smi_mut_feathers <- mutate(mut_feathers,new_smi=scaled_mass(mass,tarsus,b))  
smi_mut_feathers
# smi values look similar
```
#### Compare SMI next to each other
```{r}
select(smi_mut_feathers,smi,new_smi)
# YAY!!!! 
```
***
#### SUmmarise basal and stress corticosterone by getting mean and std_err
```{r}
sum_feathers <- group_by(smi_mut_feathers,site)
feathers_sum <- summarise(sum_feathers,
    basal_vector <- mean(basal_cort,na.rm=TRUE),
    stress_vector <- mean(stress_cort,na.rm=TRUE),
    std_err(basal_vector),
    std_err(stress_vector),
    .groups='drop')
feathers_sum
```
```{r}
std_err(basal_vector)
```








