---
title: "HW 14: Maps"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```
***
### Part 1: Outline maps
***
#### U.S. map
```{r}
# Outline of the 48 contiguous states
usa <-map_data(map="state")
str(usa)
# Outline of just Missouri
us_missouri <- map_data(map="state",region="missouri")
# Outline of Alabama, Florida, Mississippi,and North Carolina
us_other <- subset(x=usa,subset=region%in%c("alabama","florida","mississippi","north carolina"))
# Case matters with region =
```
#### Use us_other to lay first map layer
```{r}
ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")
# Interesting...
```
#### Now add Missouri to previous plot (us_missouri)
```{r}
ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")+
  geom_polygon(data=us_missouri,aes(x=long,y=lat,group=group),color="black")
# Plot appears stretched...but otherwise its ok.  Probably has to do with view.  
```
#### Add the 48 states plot to the previous two layers (usa)
```{r}
ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")+
  geom_polygon(data=us_missouri,aes(x=long,y=lat,group=group),color="black")+
  geom_polygon(data=usa,aes(x=long,y=lat,group=group),fill=NA,color="black")
# Way cool, much wow
```
#### save plot to us_map, add theme_map(),use coord_fixed of 1.3
```{r}
us_map <- ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")+
  geom_polygon(data=us_missouri,aes(x=long,y=lat,group=group),color="black")+
  geom_polygon(data=usa,aes(x=long,y=lat,group=group),fill=NA,color="black")+
  theme_map()+
  coord_fixed(1.3)
us_map
# removed grid background theme.  I like color though.  Hope we learn color.  
```
***
#### Missouri map: Highlight Scott County
```{r}
missouri <- map_data("county","missouri")
mo_scott <- subset(missouri,subregion%in%c("scott"))
# Saved missouri variable and then added it to the subset mo_scott variable
mo_map <- ggplot()+
  geom_polygon(data=missouri,aes(x=long,y=lat,group=group),fill=NA,color="black")+
  geom_polygon(data=mo_scott,aes(x=long,y=lat,group=group),fill="black")+
  theme_map()+
  coord_fixed(1.3)
mo_map
# Where da color at.  
```
***
#### Patchwork for us_map and mo_map
```{r}
us_map+mo_map+plot_layout(nrow=1,widths=c(1.5,1))
# I like it.  Much easier than yesterday
```
***
#### Utah map
```{r}
utah <- map_data("county","utah")
ut_washington <- subset(utah,subregion%in% ("washington"))
# Heard that St. George has amazing training grounds, I really want to go
ut_weber <- subset(utah,subregion%in%("weber"))
# First initial of name is William.  Ogden would be cool to visit as well.  
ut_salt_lake <- subset(utah,subregion%in%c("salt lake"))
# Middle name is Simon so S; Id like to float
# Utah is cool
ut_map <- ggplot()+
  geom_polygon(data=utah,aes(x=long,y=lat,group=group),fill=NA,color="black")+
  geom_polygon(data=ut_washington,aes(x=long,y=lat,group=group),fill="black")+
  geom_polygon(data=ut_weber,aes(x=long,y=lat,group=group),fill="grey")+
  geom_polygon(data=ut_salt_lake,aes(x=long,y=lat,group=group),fill="grey")+
  theme_map()+
  coord_fixed(1.3)
ut_map
```
***
### Part 2
***
#### Skin cancer mortality
```{r}
skin_cancer_raw <- read_delim(file=here::here("data","skin_cancer.txt"),
                              delim = " ",
                              skip=7,
                              trim_ws = TRUE)
skin_cancer_raw
```
#### Tidy
```{r}
skin_cancer <- skin_cancer_raw %>% 
  filter(State!="Wash,D.C.") %>% 
  mutate(State=case_when(State=="NorthCarolina"~"North Carolina",
                        State == "SouthCarolina" ~ "South Carolina",
    State == "NorthDakota" ~ "North Dakota",
    State == "SouthDakota" ~ "South Dakota",
    State == "NewHampshire" ~ "New Hampshire",
    State == "NewJersey" ~ "New Jersey",
    State == "NewMexico" ~ "New Mexico",
    State == "MewYork" ~ "New York", 
    State == "RhodeIsland" ~ "Rhode Island",
    State == "WestVirginia" ~ "West Virginia",
    TRUE ~ State
  ))
skin_cancer
# changing state data to proper names
```
#### Calculate mean_mort,add relative_mort column.  Create new dataset with skin_cancer_adj
```{r}
mean_mort <- mean(skin_cancer$Mort)
skin_cancer_adj <- skin_cancer %>% 
  mutate(relative_mort=Mort-mean_mort)
skin_cancer_adj
```
#### Trim Alaska,Hawaii,and territorries
```{r}
lower_48 <- c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
lower_48
```
#### Import cb_2017...500k.shp
```{r}
states <- st_read(here::here("data/shape_files","cb_2017_us_state_500k.shp"),stringsAsFactors = TRUE)
# looks ok so far, misspelled state first time around :/
```
#### add lower_48 vector to states file,use states_df
```{r}
states_df <- states %>% 
  dplyr::filter(NAME%in%lower_48)
states_df
```
#### Use STATEFP with str()
```{r}
str(states_df$STATEFP)
# looks ok so far
```
#### Capture FIPS levels from states_df; use factor for FIPS and state names 
```{r}
state_levels <- levels(states_df$STATEFP)
name_levels <- levels(states_df$NAME)
```
#### Import state FIPS csv and add previously named levels vectors
```{r}
fips <- read_csv(here::here("data","state_fips.csv")) %>% 
  dplyr::select(-abbr) %>% 
  rename(STATEFP=FIPS) %>% 
  mutate(STATEFP=factor(STATEFP,levels = state_levels))
fips
```
#### Use left join for skin_cancer,fips & states_df,skin_cancer(renamed)
```{r}
skin_cancer_all <- left_join(skin_cancer_adj,fips)
states_df_all <- left_join(states_df,skin_cancer_all)
# states_df_all is our final plotting data frame
```
```{r}
ggplot(states_df_all)+
  geom_sf(aes(fill=relative_mort))+
  scale_fill_viridis_c(name="Skin Cancer,\nRelative Mortality",
                       option="inferno")+
  coord_sf(crs=st_crs(5070))+
  theme_bw()+
  theme(legend.position = "bottom")
# This is honestly amazing!
```
***
#### Chloropleth map: your turn (Import lyme disease)
```{r}
lyme_disease_raw <- read_csv(here::here("data","lyme_disease.csv"),
       skip=1,
       na=("-"))%>%
  drop_na()
  
lyme_disease_raw
```
#### Tidy
```{r}
lyme_disease_data <- lyme_disease_raw %>% 
  mutate(`2017`=`2017_Confirmed`+`2017_Probable`) %>% 
  select(-c("2017_Confirmed","2017_Probable")) %>% 
  pivot_longer("2007":"2017",names_to="Year",values_to="Cases") %>% 
  group_by(State) %>% 
  summarise(Mean_cases=mean(Cases),
            Log_mean=log(Mean_cases),.groups="drop")
lyme_disease_data
```
#### Use left join now: Plot
```{r}
lyme_disease_all <- left_join(lyme_disease_data,fips)
lyme_states <- left_join(states_df,lyme_disease_all)
```
```{r}
lyme_states %>% 
  ggplot()+
  geom_sf(aes(fill=Log_mean))+
  scale_fill_viridis_c(name="Lyme Disease\nMean Cases",option="viridis",labels=c("<1","10","100","1000","5000"))+
  coord_sf(crs=st_crs(5070))+
  theme_bw()+
  theme(legend.position = "left")
# Got it, except for the legend being stretched when at the bottom.  so i changed it.  I like it more than the following graph
```
```{r}
lyme_states %>% 
  ggplot()+
  geom_sf(aes(fill=Log_mean))+
  scale_fill_viridis_c(name="Lyme Disease\nMean Cases",option="viridis",labels=c("<1","10","100","1000","5000"))+
  coord_sf(crs=st_crs(5070))+
  theme_bw()+
  theme(legend.position = "bottom")
# Actual graph from assignment.  
```
***
### Part 3: Dot distribution maps
***
#### Example: Distribution of _Gigantopithecus_ and _Ursus_

####set.seed & load bigfoot
```{r}
min_long <- -126
max_long <- -104
min_lat <- 31
max_lat <- 50
set.seed(9618973)

bigfoot_raw <- read_csv(here::here("data","bigfoot.csv"))
bigfoot_data <- bigfoot_raw %>% 
  filter(long>=min_long&long<=max_long,
         lat>=min_lat&lat<=max_lat) %>% 
  sample_n(300) %>% 
  mutate(name="Gigantopithecus") %>% 
  dplyr::select(name,long,lat)
bigfoot_data
```
#### Load bears and bind, find head
```{r}
bears <- read_csv(here::here("data","bears.csv"))
both_species <- bind_rows(bigfoot_data,bears)
head(both_species)
```
#### Find tail of both_species
```{r}
tail(both_species)
```
#### upload Stamen Maps
```{r}
base=get_stamenmap(bbox=c(min_long,min_lat,max_long,max_lat),
                   zoom=4,
                   maptype="terrain-background")
ggmap(base)
# Good so far
```
#### Add bigfoot_data to base
```{r}
base_bigfoot <- ggmap(base)
base_bigfoot+geom_point(data=bigfoot_data,
                aes(x=long,
                    y=lat))
```
#### Now do it for bears
```{r}
base_bears <- ggmap(base)
base_bears+geom_point(data=bears,aes(x=long,y=lat))
```
#### Now combine it with both_species
```{r}
base_species <- ggmap(base)
base_species+geom_point(data=both_species,
                    aes(x=long,
                        y=lat,
                        shape=name,
                        fill=name),
                    size=2)+
  scale_shape_manual(values=c(22:23))+
  labs(x="Longitude",
       y="Latitude",
       fill="Species",
       shape="Species")
# For some reason I type + instead of =...
# This looks ok.  
```
***
#### Set seed
```{r}
min_long <- -126
max_long <- -104
min_lat <- 31
max_lat <- 50
set.seed(1971148)
```
#### One giant code chunk
```{r}
bigfoot_raw <- read_csv(here::here("data","bigfoot.csv"))
bigfoot_data <- bigfoot_raw %>% 
  filter(long>=min_long&long<=max_long,
         lat>=min_lat&lat<=max_lat) %>% 
  sample_n(300) %>% 
  mutate(name="Gigantopithecus") %>% 
  dplyr::select(name,long,lat)
bears <- read_csv(here::here("data","bears.csv"))
both_species <- bind_rows(bigfoot_data,bears)
head(both_species)
base=get_stamenmap(bbox=c(min_long,min_lat,max_long,max_lat),
                   zoom=4,
                   maptype="terrain-background")
ggmap(base)
base_bigfoot <- ggmap(base)
base_bigfoot+geom_point(data=bigfoot_data,
                aes(x=long,
                    y=lat))
base_bears <- ggmap(base)
base_bears+geom_point(data=bears,aes(x=long,y=lat))
base_species <- ggmap(base)
base_species+geom_point(data=both_species,
                    aes(x=long,
                        y=lat,
                        shape=name,
                        fill=name),
                    size=2)+
  scale_shape_manual(values=c(22:23))+
  labs(x="Longitude",
       y="Latitude",
       fill="Species",
       shape="Species")
# It changed now, before it looked just like the previous ggplot so I add all previous codes to make it different.  
```
***
### Part 4 Bathymetry
***
#### Example: Fiji earthquakes
#### I have a Dell laptop so Windows :/.  Starting with conferting km to feet
```{r}
min_long <- -170
max_long <- 164
min_lat <- -42
max_lat <- -8

earthquakes <- quakes %>% 
  mutate(depth=depth*3280.84)
earthquakes
```
#### Run Fiji with NOAA bathyl  
```{r}
library(rgdal)
```
```{r}
memory.limit(24000)
fiji <- getNOAA.bathy(lon1=min_long,
                      lon2=max_long,
                      lat1=min_lat,
                      lat2=max_lat,
                      antimeridian=TRUE,
                      keep=TRUE)
```
```{r}
autoplot.bathy(fiji,
               geom=c("raster","contour"),
               size=0.1,
               na.rm=TRUE)
```
#### Edit
```{r}
autoplot(fiji,
         geom=c("raster","contour"),
         size=0.1,
         na.rm=TRUE)+
  scale_fill_gradient2(low="dodgerblue4",
                       mid="gainsboro",
                       high="darkgreen",
                       name="Depth (ft)")+
  labs(x="Longitude",y="Latitude")+
  theme(axis.title.y=element_text(angle=0,
                                  vjust=0.5))
```
#### Save to base_map with autoplot; add quakes data to base_map then
```{r}
base_map <- autoplot(fiji,
         geom=c("raster","contour"),
         size=0.1,
         na.rm=TRUE)+
  scale_fill_gradient2(low="dodgerblue4",
                       mid="gainsboro",
                       high="darkgreen",
                       name="Depth (ft)")+
  labs(x="Longitude",y="Latitude")+
  theme(axis.title.y=element_text(angle=0,
                                  vjust=0.5))
```
```{r}
base_map+
  geom_point(data=earthquakes,aes(x=long,y=lat))
# I feel like a real oceanographer
```
#### plot looks rough, so distinguish by magnitude
```{r}
base_map+
  geom_point(data=earthquakes,aes(x=long,y=lat,size=mag),
             alpha=0.4)
# View looks horrible with this.  
```
#### Change scale size to something more manageable and does not look blotty
```{r}
base_map+
  geom_point(data=earthquakes,aes(x=long,y=lat,size=mag),
             alpha=0.4)+
scale_size_continuous(range=c(0.5,3),name="Magnitude")
# Much better now, but view is terrible...
```
***
#### My Turn with bathymetry
#### load blennies.csv & use getNOAA.bathy(); set long, lat values
```{r}
min_long.1 <- -90
max_long.1 <- -58
min_lat.1 <- 8
max_lat.1 <- 28

blennies <- read_csv(here::here("data","blennies.csv"))
blennies
```
#### run getNOAA.bathy for carib_sea
```{r}
memory.limit(25000)
carib_sea <- getNOAA.bathy(lon1=min_long.1,
                      lon2=max_long.1,
                      lat1=min_lat.1,
                      lat2=max_lat.1,
                      antimeridian=FALSE,
                      keep=TRUE)
```
```{r}
base_map_carib <- autoplot(carib_sea,
               geom=c("raster","contour"),
               size=0.1,
               na.rm=TRUE)+
  labs(x="Longitude",
       y="Latitude")+
  scale_fill_etopo(guide=FALSE)+
  theme(axis.title.y=element_text(angle=0,
                                  vjust=0.5))
base_map_carib
# Took me a while to edit this, scrolling back and forth sometimes forces mistakes...
```
#### Add blennies to base_map_carib
```{r}
base_map_carib+
  geom_point(data=blennies,aes(x=long,y=lat,color=name,shape=name),size=2.2,fill="white")+
  scale_shape_manual(values=c(21:23))+
  labs(color="Species",shape="Species")+
  theme(legend.position = "bottom")+
  theme(legend.text=element_text(face="italic"))
# decided to keep color=name because I liked the outline of it.  
```
***
### THE END!!! ONE MORE ASSIGNMENT TO GO!!!
