---
title: "HW 14: Maps"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```
***
### Part 1: Outline maps
***
#### U.S. map
```{r}
# Outline of the 48 contiguous states
usa <-map_data(map="state")
str(usa)
# Outline of just Missouri
us_missouri <- map_data(map="state",region="missouri")
# Outline of Alabama, Florida, Mississippi,and North Carolina
us_other <- subset(x=usa,subset=region%in%c("alabama","florida","mississippi","north carolina"))
# Case matters with region =
```
#### Use us_other to lay first map layer
```{r}
ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")
# Interesting...
```
#### Now add Missouri to previous plot (us_missouri)
```{r}
ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")+
  geom_polygon(data=us_missouri,aes(x=long,y=lat,group=group),color="black")
# Plot appears stretched...but otherwise its ok.  Probably has to do with view.  
```
#### Add the 48 states plot to the previous two layers (usa)
```{r}
ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")+
  geom_polygon(data=us_missouri,aes(x=long,y=lat,group=group),color="black")+
  geom_polygon(data=usa,aes(x=long,y=lat,group=group),fill=NA,color="black")
# Way cool, much wow
```
#### save plot to us_map, add theme_map(),use coord_fixed of 1.3
```{r}
us_map <- ggplot()+
  geom_polygon(data=us_other,aes(x=long,y=lat,group=group),fill="gray")+
  geom_polygon(data=us_missouri,aes(x=long,y=lat,group=group),color="black")+
  geom_polygon(data=usa,aes(x=long,y=lat,group=group),fill=NA,color="black")+
  theme_map()+
  coord_fixed(1.3)
us_map
# removed grid background theme.  I like color though.  Hope we learn color.  
```
***
#### Missouri map: Highlight Scott County
```{r}
missouri <- map_data("county","missouri")
mo_scott <- subset(missouri,subregion%in%c("scott"))
# Saved missouri variable and then added it to the subset mo_scott variable
mo_map <- ggplot()+
  geom_polygon(data=missouri,aes(x=long,y=lat,group=group),fill=NA,color="black")+
  geom_polygon(data=mo_scott,aes(x=long,y=lat,group=group),fill="black")+
  theme_map()+
  coord_fixed(1.3)
mo_map
# Where da color at.  
```
***
#### Patchwork for us_map and mo_map
```{r}
us_map+mo_map+plot_layout(nrow=1,widths=c(1.5,1))
# I like it.  Much easier than yesterday
```
***
#### Utah map
```{r}
utah <- map_data("county","utah")
ut_washington <- subset(utah,subregion%in% ("washington"))
# Heard that St. George has amazing training grounds, I really want to go
ut_weber <- subset(utah,subregion%in%("weber"))
# First initial of name is William.  Ogden would be cool to visit as well.  
ut_salt_lake <- subset(utah,subregion%in%c("salt lake"))
# Middle name is Simon so S; Id like to float
# Utah is cool
ut_map <- ggplot()+
  geom_polygon(data=utah,aes(x=long,y=lat,group=group),fill=NA,color="black")+
  geom_polygon(data=ut_washington,aes(x=long,y=lat,group=group),fill="black")+
  geom_polygon(data=ut_weber,aes(x=long,y=lat,group=group),fill="grey")+
  geom_polygon(data=ut_salt_lake,aes(x=long,y=lat,group=group),fill="grey")+
  theme_map()+
  coord_fixed(1.3)
ut_map
```
***
### Part 2
***
#### Skin cancer mortality
```{r}
skin_cancer_raw <- read_delim(file=here::here("data","skin_cancer.txt"),
                              delim = " ",
                              skip=7,
                              trim_ws = TRUE)
skin_cancer_raw
```
#### Tidy
```{r}
skin_cancer <- skin_cancer_raw %>% 
  filter(State!="Wash,D.C.") %>% 
  mutate(State=case_when(State=="NorthCarolina"~"North Carolina",
                        State == "SouthCarolina" ~ "South Carolina",
    State == "NorthDakota" ~ "North Dakota",
    State == "SouthDakota" ~ "South Dakota",
    State == "NewHampshire" ~ "New Hampshire",
    State == "NewJersey" ~ "New Jersey",
    State == "NewMexico" ~ "New Mexico",
    State == "MewYork" ~ "New York", 
    State == "RhodeIsland" ~ "Rhode Island",
    State == "WestVirginia" ~ "West Virginia",
    TRUE ~ State
  ))
skin_cancer
# changing state data to proper names
```
#### Calculate mean_mort,add relative_mort column.  Create new dataset with skin_cancer_adj
```{r}
mean_mort <- mean(skin_cancer$Mort)
skin_cancer_adj <- skin_cancer %>% 
  mutate(relative_mort=Mort-mean_mort)
skin_cancer_adj
```
#### Trim Alaska,Hawaii,and territorries
```{r}
lower_48 <- c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
lower_48
```
#### Import cb_2017...500k.shp
```{r}
states <- st_read(here::here("data/shape_files","cb_2017_us_state_500k.shp"),stringsAsFactors = TRUE)
# looks ok so far, misspelled state first time around :/
```
#### add lower_48 vector to states file,use states_df
```{r}
states_df <- states %>% 
  dplyr::filter(NAME%in%lower_48)
states_df
```
#### Use STATEFP with str()
```{r}
str(states_df$STATEFP)
# looks ok so far
```
#### Capture FIPS levels from states_df; use factor for FIPS and state names 
```{r}
state_levels <- levels(states_df$STATEFP)
name_levels <- levels(states_df$NAME)
```
#### Import state FIPS csv and add previously named levels vectors
```{r}
fips <- read_csv(here::here("data","state_fips.csv")) %>% 
  dplyr::select(-abbr) %>% 
  rename(STATEFP=FIPS) %>% 
  mutate(STATEFP=factor(STATEFP,levels = state_levels))
fips
```
#### Use left join for skin_cancer,fips & states_df,skin_cancer(renamed)
```{r}
skin_cancer_all <- left_join(skin_cancer_adj,fips)
states_df_all <- left_join(states_df,skin_cancer_all)
# states_df_all is our final plotting data frame
```
```{r}
ggplot(states_df_all)+
  geom_sf(aes(fill=relative_mort))+
  scale_fill_viridis_c(name="Skin Cancer,\nRelative Mortality",
                       option="inferno")+
  coord_sf(crs=st_crs(5070))+
  theme_bw()+
  theme(legend.position = "bottom")
# This is honestly amazing!
```
***
#### Chloropleth map: your turn
