---
title: "HW06 Part 1: Complete the sections"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("nycflights13")
```


## Chapter 10: Tibbles

Read [R4ds Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html), sections 1-3.

### 10.1: Introduction

Load the tidyverse package. 
```{r}
library(tidyverse)
```
```{r}
library(ggplot2)
```

### 10.2: Creating tibbles

Enter your code chunks for Section 10.2 here.
Describe what each chunk code does. 
```{r}
as_tibble(iris)
#loading iris data
```
```{r}
tibble(x=1:5,y=1,z=x^2+y)
#numbers one through five = x, given x and y, calculate z
```
```{r}
tb <- tibble(':)'= "smile", ' ' = "space",'2000'="number")
#providing non-syntactic names and given their column
```
```{r}
tb
#running tb
```
```{r}
tribble(~x,~y,~z,
        #--|--|----
        "a",2,3.6,
        "b",1,8.5)
#creating a data frame given x, y, z
```


### 10.3: Tibbles vs data.frame

Enter your code chunks for Section 10.2 here.
Describe what each chunk code does. 
```{r}
tibble(a=lubridate::now()+runif(1e3)*86400,b=lubridate::today()+runif(1e3)*30,c=1:1e3,d=runif(1e3),e=sample(letters,1e3,replace=TRUE))
#honestly have no clue what this did besides give current date
```
```{r}
nycflights13::flights %>% 
  print(n=10,width=Inf)
#printing out nycflights data from 2013 showing only first 10 rows
```
```{r}
package?tibble
#trying to find more info about tibble
```
```{r}
df <- tibble(x=runif(5),y=rnorm(5))
#creating df vector
```
```{r}
df$x
#using $ to pull x variable
```
```{r}
df[["x"]]
#repeating above code chunk but using [[""]]
```
```{r}
df[[1]]
#using position 
```
```{r}
df %>% .$x
#using pipe symbol and $
```
```{r}
df %>% .[["x"]]
#using pipe symbol and [[""]]
```



### 10.4: Not required

#### Section 10.5 Questions

Answer the questions *completely.* Use code chunks, text, or both, as necessary.

**1:** How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame). Identify at least two ways to tell if an object is a tibble. *Hint:* What does `as_tibble()` do? What does `class()` do? What does `str()` do?
```{r}
data(mtcars)
```
```{r}
?mtcars
```
```{r}
print.data.frame(mtcars)
```
```{r}
as_tibble(mtcars)
```
```{r}
vignette("tibble")
#ultimately, it removes the row names from the data.  If it does not have row names, then it is a tibble.  It creates a list of columns that can be read left to right.  It is easier then to extract that variable.  
```



**2:** Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?


```{r}
df <- data.frame(abc = 1, xyz = "a")
```
```{r}
df
```
```{r}
df$x
```
```{r}
df[,"xyz"]
```
```{r}
df[, c("abc", "xyz")]
```
```{r}
tb <- tibble(`abc`="1",`xyz`="a")
```
```{r}
tb[["xyz"]]
```
```{r}
tb$xyz
#ultimately a tb has a neater appearance.  data.frame might involve more steps.  with tb, i can use $ or [[""]] and I would just need to extract the whole column name.  
```



## Chapter 11: Importing data

Read [R4ds Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html), sections 1, 2, and 5.

### 11.1 Introduction

Nothing to do here unless you took a break and need to reload `tidyverse`.

### 11.2 Getting started.

Do *not* run the first code chunk of this section, which begins with `heights <- read_csv("data/heights.csv")`. You do not have that data file so the code will not run.

Enter and run the remaining chunks in this section.

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
#creating columns a,b,c and attached numeric values
```
```{r}
read_csv("The first line of metadata
         The second line of metadata
         x,y,z
         1,2,3",skip=2)
# ignoring the first two lines of metadata by using skip=2
```
```{r}
read_csv("#A comment I want to skip
         x,y,z
         1,2,3",comment="#")
#using comment ="#" to skip first comment
```
```{r}
read_csv("1,2,3\n4,5,6",col_names = FALSE)
# using col_names = FALSE to have X1-X3 be column heads
```
```{r}
read_csv("1,2,3\n4,5,6",col_names = c("x","y","z"))
# using col_names to create a column character vector
```
```{r}
read_csv("a,b,c\n1,2,.",na=".")
# adding na to the first row, column c
```
```{r}

```



#### 11.2 Questions

**1:** What function would you use to read a file where fields were separated with "|"?
```{r}
?read_delim
```
```{r}
read_delim("a|b\n1.0|2.0", delim = "|")
#i would use read_delim, found in Examples: #File Types
```


**2:** (This question is modified from the text.) Finish the two lines of `read_delim` code so that the first one would read a comma-separated file and the second would read a tab-separated file. You only need to worry about the delimiter. Do not worry about other arguments. Replace the dots in each line with the rest of your code. 

# Comma-separated
`file <- read_delim("file.csv", delim=",")`



# Tab-separated
`file <- read_delim("file.csv", delim="|")`


**3:** What are the two most important arguments to `read_fwf()`? Why?
```{r}
?read_fwf
#specific to this function is col_types, col_positions (fwf_: empty/widths/positions) 
```



**4:** Skip this question


**5: ** Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
#this is missing c or another column variable name/letter
read_csv("a,b,c\n1,2\n1,2,3,4")
#\n adds a new line, but it is missing more row data, second \n has too much data for the columns specified
read_csv("a,b\n\"1")
# column names were specified, \n introduced new row but another \"1" screwed up what the row data was supposed to be
read_csv("a,b\n1,2\na,b")
#why do we have 2nd row values of a, b if we already have those column values
read_csv("a;b\n1;3")
#use read_csv2 for semi-colons
```

### 11.3 and 11.4: Not required

### 11.5: Writing to a file

Just read this section. You may find it helpful in the future to save a data file to your hard drive. It is basically the same format as reading a file, except that you must specify the data object to save, in addition to the path and file name.

### 11.6 Not required

## Chapter 18: Pipes

Read [R4ds Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html), sections 1-3.

Nothing to do otherwise for this chapter. Is this easy or what?


**Note:** Trying using pipes for all of the remaining examples. That will help you understand them.

## Chapter 12: Tidy Data

Read [R4ds Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html), sections 1-3, 7. 

### 12.1 Introduction

Nothing to do here unless you took a break and need to reload the `tidyverse.`
```{r}
library(tidyverse)
```


### 12.2 Tidy data

Study Figure 12.1 and relate the diagram to the three rules listed just above them. Relate that back to the example I gave you in the notes. Bear this in mind as you make data tidy in the second part of this assignment.

You do not have to run any of the examples in this section.

### 12.3

Read and run the examples through section 12.3.1 (gathering), including the example with `left_join()`. We'll cover joins later.
```{r}
table4a
```
```{r}
table4a %>% 
  pivot_longer(c(`1999`,`2000`),names_to="year",values_to="cases")
```
```{r}
?table4b 
```
```{r}
table4b
```
```{r}
table4b %>% 
  pivot_longer(c(`1999`,`2000`),names_to="year",values_to="population")
```
```{r}
tidy4a <- table4a %>% 
  pivot_longer(c(`1999`,`2000`),names_to="year",values_to="cases")
```
```{r}
tidy4b <- table4b %>% 
  pivot_longer(c(`1999`,`2000`),names_to="year",values_to="population")
```
```{r}
left_join(tidy4a,tidy4b)
```



#### 12.3 Questions

**2:** Why does this code fail? Fix it so it works.

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
# added in backticks for years 1999 and 2000.  Can run the code now.  
```

That is all for Chapter 12. On to the last chapter.


## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data Transformation](https://r4ds.had.co.nz/transform.html), sections 1-4.

Time to [get small.](https://www.youtube.com/watch?v=GOrdzCHnpw4) 

### 5.1: Introduction

Load the necessary libraries. As usual, type the examples into and run the code chunks.
```{r}
library(tidyverse)
```
```{r}
library(nycflights13)
```
```{r}
flights
```
```{r}
View(flights)
# this is cool
```



### 5.2: Filter rows with `filter()`

Study Figure 5.1 carefully. Once you learn the `&`, `|`, and `!` logic, you will find them to be very powerful tools.
```{r}
filter(flights,month==1,day==1)
# select January 1 flights
```
```{r}
jan1 <- filter(flights,month==1,day==1)
# creating jan1 vector
```
```{r}
jan1
```
```{r}
(dec25 <- filter(flights,month==12,day==25))
# creating and running dec25
```
```{r}
filter(flights,month==11|month==12)
# using | (or) feature
```
```{r}
nov_dec <- filter(flights,month %in% c(11,12))
# using %in% feature with nov_dec vector
```
```{r}
filter(flights,arr_delay<=120,dep_delay<=120)
# (flights delayed by less than 120 minutes)
```
```{r}
filter(flights,!(arr_delay>120|dep_delay>120))
# using !(|) argument
```
```{r}
NA>5
```
```{r}
10 == NA
```


#### 5.2 Questions

**1.1:** Find all flights with a delay of 2 hours or more.
```{r}
filter(flights,arr_delay>=120,dep_delay>=120)
# or
filter(flights,!(arr_delay<120|dep_delay<120))
# using both arguments to solve question
```


**1.2:** Flew to Houston (IAH or HOU)
```{r}
?flights
```
```{r}
filter(flights,dest%in% c("HOU","IAH"))
# highlighting flights to HOU or IAH airports
```


**1.3:** Were operated by United (UA), American (AA), or Delta (DL).
```{r}
filter(flights,carrier%in% c("UA","AA","DL"))
# highlighting airlines UA,AA,DL
```
```{r}
View(filter(flights,carrier%in% c("UA","AA","DL")))
```


**1.4:** Departed in summer (July, August, and September).
```{r}
filter(flights,month %in% c(7:9))
# highlighting months 7:9 flights
```


**1.5:** Arrived more than two hours late, but didn’t leave late.
```{r}
filter(flights,arr_delay>120)
```


**1.6:** Were delayed by at least an hour, but made up over 30 minutes in flight. This is a tricky one. Do your best.
```{r}
filter(flights,dep_delay>=60,arr_delay<30)
# I think this is how you do it?
```


**1.7:** Departed between midnight and 6am (inclusive)
```{r}
filter(flights,dep_time %in% c(0000:0600))
# Selecting flights between 0:00 and 6 am
```


**2:** Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
?between
# it is used instead of x >=left & x<=right
```
```{r}
filter(flights,between(dep_time,000,0600))
# i could use this code chunk to answer Q5.1.7
```



**3:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
filter(flights,is.na(dep_time))
# selecting dep_time that ended up being NA.  probably flight was ground due to weather
```


**4:** Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)
```{r}
?is.na
```
```{r}
NA^0
# any unknown or known number raised to the power of zero will always equal 1.  It does not matter.  thus NA^0 works as a code chunk
```
```{r}
NA|TRUE
```
```{r}
FALSE & NA
```
```{r}
NA*0
# ultimately, Boolean operators will result in working code chunks.  BUt doing actual math (==,+,-,*,etc.) will result in NA as the answer
```


**Note:** For some context, see [this thread](https://blog.revolutionanalytics.com/2016/07/understanding-na-in-r.html)


### 5.3 Arrange with `arrange()`
```{r}
arrange(flights,year,month,day)
```
```{r}
filter(flights)
# there both the same thing?  I guess we could have used other column names
```
```{r}
arrange(flights,desc(dep_delay))
# sorting by desc values for dep_delay
```


#### 5.3 Questions

**1:** How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()). **Note:** This one should still have the earliest departure dates after the `NA`s. *Hint:* What does `desc()` do?
```{r}
arrange(filter(flights,is.na(dep_time)))
# i guess this is how you do it?  
```
```{r}
?desc
```
```{r}
filter(flights,is.na(dep_delay))
# i can do it with filter? 
```



**2:** Sort flights to find the most delayed flights. Find the flights that left earliest. 

This question is asking for the flights that were most delayed (left latest after scheduled departure time) and least delayed (left ahead of scheduled time).
```{r}
arrange(flights,desc(dep_delay))
```
```{r}
arrange(flights,dep_delay)
```



**3:** Sort flights to find the fastest flights. Interpret fastest to mean shortest time in the air.
```{r}
arrange(flights,air_time)
# sorting by fastest air time (lower time the better)
```



*Optional challenge:* fastest flight could refer to fastest air speed. Speed is measured in miles per hour but time is minutes. Arrange the data by fastest air speed.
```{r}
arrange(flights,desc(distance/air_time))
```


**4:** Which flights travelled the longest? Which travelled the shortest?
```{r}
arrange(flights,desc(distance))
```
```{r}
arrange(flights,distance)
# sorting by longest distance and then shortest distance
```



### 5.4 Select columns with `select()`
```{r}
select(flights,year,month,day)
# highlighting year month day columns
```
```{r}
select(flights,year:day)
# same as previous code chunk but using colon
```
```{r}
select(flights,-(year:day))
# removing columns year,month,day
```
```{r}
?select
```

#### 5.4 Questions

**1:** Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights. Find at least three ways.
```{r}
select(flights,dep_delay, dep_time,arr_time,arr_delay)
```
```{r}
select(flights,starts_with("arr")|starts_with("dep"))
```
```{r}
select(flights,c(4,6,7,9))
```
```{r}
select(flights,ends_with("delay")|ends_with("time"))
```


**2:** What happens if you include the name of a variable multiple times in a `select()` call?
```{r}
select(flights,arr_time,arr_time,arr_time)
# it still works by only selecting that column.  Youre wasting time by needless typing though
```



**3:** What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?
```{r}
?one_of
# it is not as precise as any_of or all_of.  But in the following code chunk with vars vector, i was able to highlight only the five column variables
```
```{r}
flights %>% select(one_of(vars <- c("year", "month", "day", "dep_delay", "arr_delay")))
```


`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`



**4:** Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

`select(flights, contains("TIME"))`
```{r}
select(flights, contains("TIME"))
# not really, it would be similar to ends_with function as time is the suffix.  TIME is a literal word string so it made sense to use it with contains function
```



###THE END


