---
title: "HW 13: Strings and things"
author: "William Zoughaib"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
#### Load packages
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```
***
### Part 1
#### Import the beta_gliadin data.  Then change from many strings to one.
```{r}
beta_gliadin <- read_lines(here("data","beta_gliadin.txt"),skip=1)
# skip=1 skips first row
DNA <- beta_gliadin %>% 
  str_replace_all("\\d", "") %>% 
  str_replace_all("\\s", "") %>% 
  str_c(collapse="") %>% 
  str_to_upper()
DNA
# \\d equals a digit; \\s equals space.  I removed all lines into one giant line.  Then I combined 
```
#### Find total length of DNA data
```{r}
str_length(DNA)
# confirmed
```
#### Counting number of IUPAC letters using str_count.  Then count start codon of ATG, and stop codon of either TGA or TAA
```{r}
str_count(DNA,"[^A.C.T.G]")
# answer is 0 as I am excluding A,C,G,T from the count
str_count(DNA,"ATG")
# answer is 66
str_count(DNA, "TGA|TAA")
# first answer I got was 0 because I had spaces before and after | .  Then I deleted them so real answer is 99.  Always double check what you put in the "|"
```
#### Count number of CCAAT (CAAT box) and number of CTATAA(TATA box).  Make sure to assign them to caat_box and tata_box
```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"
str_count(DNA,c(caat_box,tata_box))
# officially 4 of caat_box and 1 of tata_box
```
#### Find caat_seq and tata_seq for CCAAT and CTATAA respectively.  Each of these ends with ATG.  
```{r}
caat_seq <- "CCAAT.{100,150}ATG"
caat_seq
str_count(DNA,caat_seq)
tata_seq <- "CTATAA.{50,100}ATG"
tata_seq
str_count(DNA,tata_seq)
# I correctly understood that .= filler characters.  But I did not know how to include the 100,150 and 50,100 respectively.  
```
#### Extract caat_seq and save in caat_seqs_all.  Use simplify=TRUE.  Find length.  
```{r}
caat_seqs_all <- str_extract_all(DNA,caat_seq,simplify = TRUE)
caat_seqs_all
str_length(caat_seqs_all)
# Difference in length of both strands is 9 n.t.
```
#### Print promotor and determine length.  Promotor will have tata_box in the caat_seqs_all
```{r}
str_which(caat_seqs_all,tata_box)
promotor <- str_sub(caat_seqs_all)
promotor[1]
str_length(promotor[1])
# This took a while.  String is 139 n.t. and i think it matches.  
```
***
### Part 2
#### Import ants.csv
```{r}
ants_data <- read_csv(here("data","ants.csv")) 
ants_data
ants_data[1:6,]
# imported data and selected first 6 rows.  I did not save it as it will skew graphing data.  
```
#### Wrangle ants_data
```{r}
ants_tidy <- ants_data %>% 
  pivot_longer(names_to="species",values_to="abundance",matches("([A-Z]).+\\..+",ignore.case = FALSE)) %>% 
  select(!(season:sample))
ants_tidy
# This is a hard assignment.  I dont even know if this is correct.  I for sure did remove season:sample columns
```


```{r}
ants_mut <- ants_tidy %>% mutate(species=str_replace(species,"([A-Z].{3}).+\\.([a-z]{1,4}).*","\\1\\2")) %>% 
  mutate(year = factor(year,ordered=TRUE,levels=c(2010,2011))) %>%
   group_by(year, species) %>%
   summarise(total = sum(abundance))
ants_mut
# so apparently all i needed was a dang * to add...otherwise everything was right... what wasted time...
```
#### Plot ants_data
```{r}
ants_mut %>% 
  ggplot(aes(x=reorder(species,total),y=total,group=species))+
  geom_line(color="grey")+
  geom_point(aes(shape=year),size=2)+
  coord_flip()+
labs(x=NULL,y="Total Abundance",shape="Year")
theme_minimal()  
# I guess I got it right...
```
***
### Part 3
#### Import flowers_rough using read_csv
```{r}
flowers_rough <- read_csv(here("data","flower_size.csv")) 
flowers_rough 
```
#### Wrangle
```{r}
flowers_data <- flowers_rough %>% 
  drop_na("Flower number") %>% 
  select("species"=Species,"cor_length"=`Corolla length (cm)`,"throat_width"=`Corolla width throat (cm)`) %>% 
  separate(species,into=c("genus","species"),sep = " ") %>%   group_by(genus,species) %>% 
  summarise(mean_length=mean(cor_length,na.rm = TRUE),
            mean_width=mean(throat_width,na.rm = TRUE),
            max_length=max(cor_length,na.rm = TRUE),
            max_width=max(throat_width,na.rm = TRUE),.groups='drop') 
flowers_data
# rename columns using select, drop_na(),seperate Genus Species, then of course use group_by and summarise together.  
```
#### Import second data set into pollinators_raw
```{r}
pollinators_raw <- read_csv(here("data","flower_pollinators.csv"))
pollinators_raw
```
#### Wrangle
```{r}
pollinators_data <- pollinators_raw %>% 
  select(-(Color|starts_with("Source"))) %>% 
  drop_na("Pollinator") %>% 
  separate(Species,into = c("genus","species"),sep = " ") %>% 
  separate(Pollinator,into="Syndrome",sep=",| ",extra="drop")
pollinators_data
# This looks ok. I dropped NAs, removed color and Source_ columns.  Separated (need to spell that correctly) Species into genus and species.  Pollinator into Syndrome.  By sep=",| " i am removing , and spaces from the carriers.  Nice and Tidy
```
#### Use inner join.  
```{r}
phlox <- inner_join(flowers_data,pollinators_data,by=c("genus","species"))
phlox
# Had to mess around with this for a bit.  NAs kept showing up.  Had to ultimately add na.rm to all the summarise arguments.  
```
####Plot
```{r}
phlox %>% 
  ggplot(aes(x=mean_length,y=mean_width,color=Syndrome))+
  geom_point()+
  labs(x="Mean length",y="Mean width")
# First plot looks ok.  I have 11 carriers.  adjusted x and y axis names
```
#### Drop bee and beetle and convert into bees and beetles respectively
```{r}
phlox_mut <- phlox %>% 
  mutate(Syndrome=str_replace(Syndrome,"(^be.*e$)", "\\1s"))
phlox_mut
# I presume this is correct?
```
#### Replot; Count
```{r}
phlox_mut %>% 
  ggplot(aes(x=mean_length,y=mean_width,color=Syndrome))+
  geom_point()+
  labs(x="Mean length",y="Mean width")
# I now have 9 pollen carriers
phlox_mut %>% 
  group_by(Syndrome) %>% 
  count(Syndrome)
```
#### Lumb beetles and flies as other insects: replot.  
```{r}
mut_phlox <- phlox_mut %>% 
  mutate(Syndrome=str_replace(Syndrome,".+l.?es","other insects"))
mut_phlox %>% 
  ggplot(aes(x=mean_length,y=mean_width,color=Syndrome))+
  geom_point()+
  labs(x="Mean length",y="Mean width")
# Similar and now 8 carriers
```
#### Use mutate w/ case_when to create syndrome_group column.  autogamous becomes Autogamous, bat to Mammal, hummingbird to Bird, everthing else to Insect.  Then plot
```{r}
phlox_data <- mut_phlox %>% 
  mutate(syndrome_group=case_when(Syndrome=="autogamous"~"Autogamous",Syndrome=="bat"~"Mammal",Syndrome=="hummingbird"~"Bird",TRUE~"Insect"))
phlox_data
# Dang, this is hard.  But i created a syndrome_group column.  Syndrome names will match whatever is next to it in syndrome_group
phlox_data %>% 
  ggplot(aes(x=mean_length,y=mean_width,color=syndrome_group))+
  geom_point()+
  labs(x="Mean length",y="Mean width",color="Syndrome")
# It was never mentioned in the directions but I presume we are use color=syndrome_group now.  It looks correct if so.  
```
#### GGhighlight
```{r}
phlox_data %>% 
  ggplot(aes(x=mean_length,y=mean_width,color=syndrome_group))+
  geom_point()+
  labs(x="Mean length",y="Mean width",color="Syndrome")+
  gghighlight(syndrome_group=="Mammal"|syndrome_group=="Bird",use_direct_label = FALSE)
# Interesting.  I forgot to do syndrome_group=="Bird" with |.   
```
#### Improvements: colors and themes
```{r}
phlox_data %>% 
  ggplot(aes(x=mean_length,y=mean_width,color=syndrome_group))+
  geom_point()+
  scale_color_brewer(type = "qual", palette = "Dark2") +
  labs(x = "Mean Length (cm)",
       y = "Mean Width (cm)",
       color = "Syndrome")
# using scale_color_brewer
```
#### Using SEMO colors (who woulda thunk it)
```{r}
semo_palette <- c("#C8102E", "#DBE245", "#71C5E8", "#B07C57")

phlox_data %>% 
  ggplot() +
  geom_point(aes(x = mean_length,
                 y = mean_width,
                 shape = syndrome_group,
                 fill = syndrome_group)) +
  scale_shape_manual(values = c(22:25)) +
  scale_fill_manual(values = semo_palette) +
  labs(x = "Mean Length (cm)",
       y = "Mean Width (cm)",
       shape = "Syndrome",
       fill = "Syndrome")
# copied and pasted
```
#### changing the lettering...
```{r}
phlox_data %>% 
  ggplot() +
  geom_point(aes(x = mean_length,
                 y = mean_width,
                 shape = syndrome_group,
                 fill = syndrome_group),
             size = 2) +
  labs(shape = "Syndrome",
       fill = "Syndrome",
       x = "Mean Length (cm)",
       y = "Mean Width (cm)") +
  theme_tufte() +
  scale_shape_manual(values = c(22:25)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))
# There are over 20-25 different themes?  I presume this means font...
```
```{r}
phlox_data %>% 
  ggplot() +
  geom_point(aes(x = mean_length,
                 y = mean_width,
                 shape = syndrome_group,
                 fill = syndrome_group),
             size = 2) +
  labs(shape = "Syndrome",
       fill = "Syndrome",
       x = "Mean Length (cm)",
       y = "Mean Width (cm)") +
  theme_economist () +
  scale_shape_manual(values = c(22:25)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))
# playing around but this is really ugly...
```
***
#### Part 4: 4 plots
```{r}
phlox_data %>% 
  ggplot(aes(x=max_length,y=max_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="Syndrome",fill="Syndrome")+
  theme_classic()+
  scale_shape_manual(values=c(22:25))+
  scale_fill_brewer(palette="Dark2")
```
```{r}
phlox_data %>% 
  ggplot(aes(x=max_length,y=max_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="Syndrome",fill="Syndrome")+
  theme_economist()+
  scale_shape_manual(values=c(22:25))+
  scale_fill_brewer(palette="Dark2")
```
```{r}
library(viridis)
```
```{r}
phlox_data %>% 
  ggplot(aes(x=max_length,y=max_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="Syndrome",fill="Syndrome")+
  theme_bw()+
  scale_shape_manual(values=c(22:25))+
  scale_fill_viridis_d(option = "magma")
```
```{r}
library(viridis)
```
```{r}
phlox_data %>% 
  ggplot(aes(x=max_length,y=max_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="Syndrome",fill="Syndrome")+
  theme_tufte ()+
  scale_shape_manual(values=c(22:25))+
  scale_fill_viridis_d(option = "viridis")
```
***
#### THE END FINALLY HOURS AND HOURS LATER




